---
title: "Adding Data Sets"
output:
  BiocStyle::html_document:
  toc: true
  toc_float: true
number_section: true
self_contained: true
titlecaps: true
vignette: >
  %\VignetteIndexEntry{Adding Data Sets}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{css, echo=FALSE}
.scroll250 {
  max-height: 250px;
  overflow-y: auto;
  background-color: inherit;
}
```


# Introduction

This package is a data repository for single cell multiome data within `ExperimentHub`. It also provides a framework to add more multiome data with minimal effort but the process cannot be fully automated. This vignette will guide you.

Briefly, data set will be disassembled and saved in a hdf5 file, from which individual experiments can be drawn using its own accessor function. You must provide the data set in hdf5 format, its metadata, its documentation, and finally the accessor function that will retrieve the data from `ExperimentHub`.



<br><br><br>

# Developer Mode

Adding data involves updating the package and as such it must be done in "developer mode". The developer mode allows access to additional tools such as documentation templates.

To work in developer mode, you must first clone the package repository with `git`. Create a branch from `master` and work on that.

Start an R session in the package directory (e.g. by opening the RStudio project in RStudio) and load all the functions. This is necessary for the R engine to temporarily identify your working directory as the package installation directory, and to expose the internal functions that you will be using.

```{r devtoolsAvailable, include = FALSE}
devtoolsAvailable <- requireNamespace("devtools", quietly = TRUE)

if (devtoolsAvailable) {
    # attach development version of the package
    devtools::load_all()
}

```
```{r devtools, eval = FALSE}
devtools::load_all()
```



<br><br><br>

# Saving Your Data Set

Currently only `MultiAssayExperiment` objects are supported. Experiments must be objects that inherit from `SummarizedExperiment` and will usually be `SingleCellExperiment`, hence full support is provided for the latter and their slots (`reducedDims` and `altExps`).

If you start with an ArchR project, you can use `archr2MAE` to convert to it a `MultiAssayExperiment`.
```{r archr2MAE, eval = FALSE}
# convert ArchR project to MAE
MAE <- archr2MAE("<ARCHR_PATH>")
```


Use `saveMAE` to save your data set to the `R/data` directory. You can use `testFile` to validate that your data set can be reconstructed.

```{r saving, eval = devtoolsAvailable, class.output = "scroll250", class.echo = "scroll250"}
# construct a dummy data set
mae <- dummyMAE()
mae

# name the file to save to
fileName <- tempfile(fileext = ".h5")

# save data set
saveMAE(mae, fileName)

# test reloading
testFile(fileName)
```

For a detailed explanation of the process see `?saveMAE`.



<br><br><br>

# Creating Metadata

Run `makeMakeMetadata("<DATASET_NAME>")` to create a script to store your data set's metadata. It will be called `inst/scripts/make-metadata-<DATASET_NAME>.R`. Metadata must be a 1-row data frame with specific columns and values must be character strings (some fields allow character vectors). See `inst/scripts/make-metadata.R` for more information and `inst/scripts/make-metadata-prostateENZ.R` for an example.

The file also stores metadata that will be returned by `listDatasets`. Likewise, this must be a 1-row data frame and values must be character strings.


```{r making metadata, eval = FALSE}
makeMakeMetadata("dataset")
```

Once your `make-metadata` file is ready, build the metadata by running the `inst/scripts/make-metadata.R` script. This must run without errors. If successful,
all the datasets will be captured in `inst/extdata/manifest.csv` and `inst/extdata/metadata.csv`

```{r build metadata, eval = FALSE}
# this will add your metadata
source(system.file("scripts", "make-metadata.R", package = "scMultiome"))
```

Subsequently, validate the correctness of the metadata by the running `makeExperimentHubMetadata` function. This call must also run without errors. It will return an `ExperimentHub` object that will display your metadata in the form that the end users will see it.


```{r validation, eval = FALSE, class.output = "scroll250"}
# this will validate your metadata (not shown)
ExperimentHubData::makeExperimentHubMetadata(dirname(system.file(package = "scMultiome")))
```



<br><br><br>

# Documenting Your Data Set

Every data set needs a help page to describe the source data, references, and whatever else you deem necessary.
Since every data set is accessed by its own accessor function, the `R` file that defines the function is the natural place to place the data set documentation. The package framework is constructed such that accessor functions are extremely simple and you can basically copy the original one (`prostateENZ`) and most of its documentation.

The process of obtaining the data set should also be described.

First, run `makeMakeData("<DATASET_NAME>")`. This will create an Rmarkdown file called `inst/scripts/make-data-<DATASET_NAME>.Rmd`. Use it to describe how your data set was obtained. This doesn't have to be a working script, just a report. Pseudocode is acceptable. Note that code evaluation has been disabled so that you can copy your actual code and the lengthy ArchR analysis does not run again.

```{r making, eval = FALSE}
makeMakeData("dataset")
```

Second, run `makeR("<DATASET_NAME>")`. This will create an R file called `<DATASET_NAME>.R`, which will provide access and help for your data set. This file will quote the accompanying Rmd file created above. This way the R file itself is more concise and easier to edit. Adjust the file accordingly:

1. Give the file the same Title was used in the `make-metadata-<>.R` file.
2. Add a Description section.
3. Document any arguments other than `experiments` and `metadata`.
4. Describe the format of your `MultiAssayExperiment`.
5. Cite appropriate references.
6. Make sure the default value of the `experiments` argument reflects the experiment names in your data set.
7. If you want to restore custom classes to your experiments, add converting functions here. They do not require documentation or exporting.


```{r making2, eval = FALSE}
makeR("dataset")
```

Once the files are ready, build the documentation, and `?scMultiome` and `?<DATASET_NAME>` to review it.
```{r documentation, eval = FALSE}
# build documentation
devtools::document()

# view package man page
?scMultiome

# view your data set man page
help("dataset")
```



<br><br><br>

# Updating Package Files

Edit the `DESCRIPTION` file to update package metadata: add yourself as package author and package version.

```{r DESCRIPTION, eval = FALSE}
desc::desc_add_author("<GIVEN_NAME>", "<FAMILY_NAME>", "<EMAIL>", role = "aut")
desc::desc_bump_version("minor")
```

Push your changes to git and create a merge request. Once the request is approved and merged to the `master` branch, the package maintainer will take over and move on to the next stage.



<br><br><br>

# Uploading Your Data Set

The final step of the process is to place the data set in Bioconductor's data bucket. This can only be done with Bioconductor's knowledge and blessing.

Bioconductor will be notified that a `scMultiome` update is coming by email at `hubs@bioconductor.org`. They will issue a temporary SAS token for the Bioconductor data bucket. The data set will be placed in the Bioconductor staging directory with `uploadFile` and Bioconductor will be notified again that the upload is ready. They will receive a link to the package repository, update metadata in `ExperimentHub` and finalize the process.

```{r upload, eval = FALSE}
uploadFile(file = fileName, sasToken = "<SAS_TOKEN>")
```

<br>

Consult [this vignette](https://www.bioconductor.org/packages/release/bioc/vignettes/HubPub/inst/doc/CreateAHubPackage.html#additional-resources-to-existing-hub-package) for current Bioconductor requirements.



<br><br><br>

# Congratulations

If you found any of this vignette or the process confusing, we would welcome feedback and gladly add more clarifications. Please email the package maintainer, `r desc::desc_get_maintainer()`.



<br><br><br>
