---
  title: "Single-cell analysis of samples from healthy human colon"
author: "Tomasz Włodarczyk"
date: 2022-09-21
output:
  BiocStyle::html_document:
  titlecaps: false
toc_float: true
---
  
```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```
scATAC data was downloaded from Gene Expression Omnibus
(acc. no. [GSE165659](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE165659))
and analyzed with SingleCell ATAC - 10X pipeline v2.0.0 \cr
scRNAseq data in form of Seurat objects was downloaded from  https://drive.google.com/drive/folders/12j9ufV1L0uWbUlab-VoXRznDLKDO7PQ. In case
of future change in the data storage location, it will be updated in the readme file in project's Github repository (https://github.com/winstonbecker/scCRC_continuum) \cr

Downstream analysis was performed with the \code{ArchR} package v. 1.0.2:


```{r}
library(ArchR)
library(parallel)

catlas_files <- <FRAGMENT_FILES>
outputDir <- <OUTPUT_DIRECTORY>
arrow_files <- createArrowFiles(inputFiles = catlas_files, sampleNames = <SAMPLE_NAMES>)
doubScores <- addDoubletScores(input = arrow_files)

# cerate ArchR project
project <- ArchRProject(arrow_files, outputDirectory = outputDir)

# filtering out doublet cells
project <- filterDoublets(project)

# add Iterative Latent Semantic Indexing reduced-dimensionality space
project <- addIterativeLSI(ArchRProj = project, useMatrix = "TileMatrix", 
                           name = "IterativeLSI", clusterParams = list(resolution = c(0.2), 
                                                    sampleCells = 10000, n.start = 10))

# batch correction
project  <- addHarmony( ArchRProj = project, reducedDims = "IterativeLSI",
    name = "Harmony", groupBy = "Sample")

project <- addClusters( input = project, reducedDims = "IterativeLSI",
    method = "Seurat", name = "Clusters", resolution = 0.8)

# add clusters after Harmony batch correction
project <- addClusters(input = project, reducedDims = "Harmony",
    method = "Seurat", name = "Clusters_Harmony", resolution = 0.8)

# add UMAP embedding 
project <- addUMAP(ArchRProj = project, reducedDims = "IterativeLSI",
    nNeighbors = 30, minDist = 0.5, name ="UMAP_LSI")

project <- addUMAP(ArchRProj = project, reducedDims = "IterativeLSI",
    nNeighbors = 30, minDist = 0.5, name ="UMAP_Harmony")

# add column with log base 10 of the fragment numbers
project <- addCellColData(ArchRProj = project, data = log10(project$nFrags),
                                  name = "log10_nFrags", cells = project$cellNames)




# upload gene expression data
# use files downloaded from 
# https://drive.google.com/drive/folders/12j9ufV1L0uWbUlab-VoXRznDLKDO7PQ_?usp=sharing

data_files <- c("Final_scHTAN_colon_normal_epithelial_220213.rds",
                "Final_scHTAN_colon_immune_220213.rds",
                "Final_scHTAN_colon_stromal_220213.rds")

# define object names
RNAseq_se_names <- gsub(".*scHTAN_|_220213.rds", "", data_files)

# create objects as instances of SingleCellExperiment class
for (i in seq_along(RNAseq_se_names)) assign(RNAseq_se_names[i], Seurat::as.SingleCellExperiment(readRDS(data_files[i])))

# add column with cell types and disease state
for (obj in RNAseq_se_names){
    eval(parse(text = paste0("colData(", obj, ")$CellType <-", obj,"@colData@listData$CellType")))
    eval(parse(text = paste0("colData(", obj, ")$DiseaseState <-", obj,"@colData@listData$DiseaseState")))
}

# uniformize colData columns before merging
shared_cols <- purrr::map(list(colon_immune, colon_stromal, colon_normal_epithelial), colData) %>%
    purrr::map(colnames) %>%
    purrr::reduce(intersect)

# remove reducedDims since their column names differ across objects
for (obj in RNAseq_se_names){
    eval(parse(text = paste0(obj, "@colData <- ", obj, "@colData[,colnames(", obj, "@colData) %in% shared_cols]")))
    eval(parse(text = paste0("SingleCellExperiment::reducedDim(", obj, ") <- NULL")))
    eval(parse(text = paste0(obj, "@int_colData@listData <- list()")))
}

# merge RNAseq data objects

colon_RNAseq <- cbind(colon_immune, colon_normal_epithelial, colon_stromal)


RNA_se <- SummarizedExperiment(assay = list(counts = as(assay(colon_RNAseq, "counts"), "dgCMatrix")),
                               colData = colData(colon_RNAseq), rowData = rowData(colon_RNAseq))


# select samples from healthy donors (no cancer)
RNA_se <- RNA_se[,colData(RNA_se)$DiseaseState == "Normal"]

# RNA integration
project <- addGeneIntegrationMatrix(
    ArchRProj = project,
    useMatrix = "GeneScoreMatrix",
    reducedDims = "IterativeLSI",
    seRNA = RNA_se,
    addToArrow = TRUE,
    groupRNA = "CellType",
    nameCell = "predicted_cell_un",
    nameGroup = "predicted_group_un",
    nameScore = "predicted_score_un")


project <- addGroupCoverages(ArchRProj = project, groupBy = "predicted_group_un")

# add pseudo-bulk replicates
## requires MACS2 installation

project <- addReproduciblePeakSet( ArchRProj = project,
    groupBy = "predicted_group_un", pathToMacs2 = <PATH_TO_MACS2>)


# LSI reduced dimensionality based on the GeneIntegrationMatrix

project <- addIterativeLSI(ArchRProj = project, clusterParams = list(resolution = 0.2,
        sampleCells = 1000, n.start = 10), saveIterations = FALSE,
    useMatrix = "GeneIntegrationMatrix", varFeatures = 2500,
    firstSelection = "variable", binarize = FALSE, name = "LSI_RNA")

# add clusters based on the new reduced-dimensionality space
project <- addClusters( input = project, reducedDims = "LSI_RNA",
    method = "Seurat", name = "Clusters_RNA", resolution = 0.8)

# add UMAP embedding
project <- addUMAP(ArchRProj = project, reducedDims = "LSI_RNA",
    nNeighbors = 30, minDist = 0.5, name ="UMAP_LSI_RNA", metric = "cosine",
    force = TRUE)

# batch correction 
project  <- addHarmony( ArchRProj = project, reducedDims = "LSI_RNA",
    name = "Harmony_RNA", groupBy = "Sample")

# UMAP embedding after batch correction
project <- addUMAP(ArchRProj = project, reducedDims = "Harmony_RNA",
    nNeighbors = 30, minDist = 0.5, name ="UMAP_LSI_RNA_Harmony", metric = "cosine",
    force = TRUE
)

# find clutsters after batch correction
project <- addClusters( input = project, reducedDims = "Harmony_RNA",
    method = "Seurat", name = "Clusters_RNA_Harmony", resolution = 0.8)


# combine reduced-dimensionality spaces produced from ATACseq and RNAseq data
project <- addCombinedDims(project, reducedDims = c("IterativeLSI", "LSI_RNA"),
                            name =  "LSI_Combined")

# add UMAP embedding
project <- addUMAP(ArchRProj = project, name = "UMAP_combined", reducedDims = "LSI_Combined",
        nNeighbors = 30, minDist = 0.5, metric = "cosine")


# find clusters in combined reduced space
project <- addClusters(input = project, reducedDims = "LSI_Combined",
                        method = "Seurat", name = "Clusters_combined",
                        resolution = 0.4)

# add information about sequence motifs recognized by known transcriptions factors
project <- addMotifAnnotations(ArchRProj = project,
                                       motifSet = "cisbp", name = "Motif")


# add background peaks to be compared against during peak variation assessement
project <- addBgdPeaks(project)

# calculate per-cell devations of motif annotations
project <- addDeviationsMatrix(project, peakAnnotation = "Motif")

# save project
saveArchRProject(project, outputDir)

# convert project into MultiAssayExperiment object 
MAE <- maw.archr::create.mae.with.multiple.sces.from.archr(outputDir, tile.sizes = 500)

saveRDS(MAE, <OUTPUT_PATH>)

```
