---
title: "LNCaP Cells Data Set"
author: "Aleksander Chlebowski"
date: 6 September 2022
output:
  BiocStyle::html_document:
    titlecaps: false
    toc_float: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```


scATAC data was downloaded from Gene Expression Omnibus
(acc. no. [GSE168667](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE168667))
and analyzed with SingleCell ATAC - 10X pipeline v2.0.0. \cr
scRNAseq data was downloaded from  Gene Expression Omnibus
(acc.no. [GSE168668](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE168668))
and analyzed with SingleCell Gene Expression Analysis - 10X pipeline v6.0.1. \cr

Downstream analysis was performed with the \code{ArchR} package:

#### 1. Initiate `ArchR` project
```{r}
# attach ArchR package
library(ArchR)

# configure ArchR
addArchRThreads(16L)
addArchRGenome("hg38")

# create arrow file from fragment files
## list fragment files
fragments <- <FRAGMENT_FILES>
## assign sample names
names(fragments) <- <SAMPLE_IDs>
## create arrows
createArrowFiles(inputFiles = fragments, sampleNames = names(fragments))

# specify output directory
outDir <- <OUTPUT_DIRECTORY>

# locate arrow files
arrows <- <ARROW_FILES>

# create ArchR project
project <- ArchRProject(arrows, outDir)

# add sample annotation
sampleNames <- c("SRR13927735", "SRR13927736", "SRR13927737", "SRR13927738")
sampleCells <- c("LNCaP", "LNCaP", "LNCaP RES-A", "LNCaP RES-B")
sampleTreatment <- c("0.1% DMSO 48h", "enzalutamide 48h", "enzalutamide", "enzalutamide")
sampleEnzalutamide <- c("sensitive", "sensitive", "resistant", "resistant")
names(sampleCells) <- names(sampleTreatment) <- names(sampleEnzalutamide) <- sampleNames
project$Cells <- sampleCells[project$Sample]
project$Treatment <- sampleTreatment[project$Sample]
project$Enzalutamide <- sampleEnzalutamide[project$Sample]
project$sampleLabels <- sampleLabels[project$Sample]
```

#### 2. Prepare RNA-seq data
```{r}
# gene expression data is analyzed with `scran.chan` package
# the result is a SingleCellExperiment object
SCE <- <scran.chan ANALYSIS>
    
# adjust for integration
rownames(SCE) <- rowData(SCE)$Symbol
assay(SCE, "counts") <- as(assay(SCE, "counts"), "dgCMatrix")
# drop duplicates
SCE <- SCE[!duplicated(rowData(SCE)$Symbol), ]
```

#### 3. Commence `ArchR` analysis
```{r}
# reduce dimensionality by iterative LSI
project <- addIterativeLSI(project, useMatrix = "TileMatrix", name = "iLSI_ATAC")

# integrate ATAC and RNAseq
## prepare grouping for constrained integration
groupMapping <- SimpleList(
    sens_NT = SimpleList(
        ATAC = project$cellNames[project$Sample == "SRR13927735"],
        RNA = grep("SRR13927739", colnames(SCE), value = TRUE)
    ),
    sens_Enz = SimpleList(
        ATAC = project$cellNames[project$Sample == "SRR13927736"],
        RNA = grep("SRR13927740", colnames(SCE), value = TRUE)
    ),
    RES_A = SimpleList(
        ATAC = project$cellNames[project$Sample == "SRR13927737"],
        RNA = grep("SRR13927741", colnames(SCE), value = TRUE)
    ),
    RES_B = SimpleList(
        ATAC = project$cellNames[project$Sample == "SRR13927738"],
        RNA = grep("SRR13927742", colnames(SCE), value = TRUE)
    )
)
## execute
project <- addGeneIntegrationMatrix(project, useMatrix = "GeneScoreMatrix",
                                    matrixName = "GeneIntegrationMatrix",
                                    reducedDims = "iLSI_ATAC", seRNA = SCE,
                                    groupATAC = "Sample", groupRNA = "Sample", groupList = groupMapping,
                                    nameCell = "predictedCell",
                                    nameGroup = "predictedGroup",
                                    nameScore = "predictedScore",
                                    addToArrow = TRUE, force = TRUE)

# add LSI for RNAseq
project <- addIterativeLSI(project, useMatrix = "GeneIntegrationMatrix", name = "iLSI_RNAseq")

# combine dim-reduced ATAC and RNAseq
project <- addCombinedDims(project, name = "iLSI_Combined", reducedDims = c("iLSI_ATAC", "iLSI_RNAseq"))

# add UMAP embedding on combined reduced dimensionality
project <- addUMAP(project, reducedDims = "iLSI_Combined", name = "UMAP_Combined", verbose = FALSE)

# impute weights (for smoother visualizations)
project <- addImputeWeights(project, reducedDims = "iLSI_Combined")

# add group coverages
## inspect available cell numbers
table(project$Sample)
project <- addGroupCoverages(project, groupBy = "Sample", minCells = 30, maxCells = 250)

# add pseudo-bulk replicates
## requires MACS2 installation
project <- addReproduciblePeakSet(project, groupBy = "Sample", pathToMacs2 = "<PATH_TO_MACS2_INSTALLATION>")

# add peak matrix
project <- addPeakMatrix(project)
getAvailableMatrices(project)

# add motif annotation
project <- addMotifAnnotations(project, motifset = "cisbp", name = "Motif")

# add background peaks
project <- addBgdPeaks(project, method = "chromVAR")

# add deviation matrix
project <- addDeviationsMatrix(project, peakAnnotation = "Motif", force = TRUE)
getAvailableMatrices(project)

# save ArchR project
saveArchRProject(project)
```

#### 4. Save results
```{r}
# convert project to MultiAssayExperiment object
MAE <- maw.archr::create.mae.with.multiple.sces.from.archr(outDir)

# inspect object
MAE

# remove unpublished class
ind <- which(names(MAE) == "TileMatrix500")
MAElim <- MultiAssayExperiment::MultiAssayExperiment(
    experiments = c(
    TileMatrix500 = as(experiments(MAE)[[ind]], "SingleCellExperiment"),
    as.list(experiments(MAE)[-ind])
))

# save object
saveMAE("inst/extdata/prostateENZ.h5")
```

\cr
